---
title: "lmmHeterosced"
author: "Geir H. Bolstad"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Weight}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
<!---  
output: 
  word_document:
    fig_caption: true
-->


<!--- 
                          NB! Build & Reload does not create a vignette 
                          use devtools::build(), or use menu for 
                          building source package
                          or devtools::build_vignettes()
-->


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(lmmHeterosced)
```

# Simulated data
```{r data}
n_cat <- 10
n <- 50
d <- data.frame(x = rnorm(n*n_cat, 0, 2), cat = rep(1:n_cat, each = n))
d$y <- rnorm(nrow(d), 10, exp(0.5*(1+0.5*d$x))) + rep(rnorm(n_cat), each = n)
with(d, plot(x,y))
```

# The model
```{r analysis}
mod <- lmmHeterosced(formula = y~1+(1|cat), heterosced_formula = ~ 1 + x, data = d)
```

# Results of Heteroscedasticity analysis
The intercept gives the log residual variance at x = 0 and the linear parameter (x) gives the change in log residual variance with x
```{r}
mod$Heterosced
```

# Plot of change in residual standard deviation with x
```{r}
with(d, plot(x,y))
x <- seq(min(d$x), max(d$x), length.out = 100)
a <- mod$Heterosced[1,1]
b <- mod$Heterosced[2,1]
lines(x, mod$Fixef[1,1] + sqrt(exp(a + b*x)), col = "red")
lines(x, mod$Fixef[1,1] - sqrt(exp(a + b*x)), col = "red")
```

# Plot of residual squares and change in variance with x
```{r}
with(d, plot(x,mod$Residuals[,1]^2, ylab = "residuals squared"))
lines(x, exp(a + b*x), col = "red")
coefGLM <- coef(glm(mod$Residuals[,1]^2 ~ d$x, family = gaussian(link = "log")))
lines(x, exp(coefGLM[1] + coefGLM[2]*x), col = "green")

```

# Plot of log residual squares and change in log variance with x
```{r}
with(d, plot(x,log(mod$Residuals[,1]^2), ylab = "log residuals squared"))
lines(x, a + b*x, col = "red")
lines(x, coefGLM[1] + coefGLM[2]*x, col = "green")
abline(lm(log(mod$Residuals[,1]^2) ~ d$x), col = "green", lty= "dashed")
```



